---
title: "p8105_hw2_amh2325"
author: "Amanda Howarth"
date: "10/2/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
library(readxl)
```
## Problem 1
## Import trash_wheels data sheet 1
```{r}
trash_wheels_data =
  read_excel(path = "./data/trash_wheel.xlsx", sheet = 1) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  select(-15, -16, -17) %>%
  mutate(sports_balls = as.integer(sports_balls))
trash_wheels_data
```
## Import precipitation_2018 data
```{r}
precipitation_2018_data = 
  read_excel(path = "./data/trash_wheel.xlsx", sheet = 6, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2018)
precipitation_2018_data
```
## Import precipitation_2017 data
```{r}
precipitation_2017_data = 
  read_excel(path = "./data/trash_wheel.xlsx", sheet = 5, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2017)
precipitation_2017_data
```
## Merging datasets to form precipiation_join data 
```{r}
precipitation_join = 
  full_join(precipitation_2018_data, precipitation_2017_data) %>%
  mutate(month = as.numeric(month), month = month.name[month])
precipitation_join
```
## Dataset summaries 
Write a paragraph about these data; you are encouraged to use inline R. Be sure to note the number of observations in both resulting datasets, and give examples of key variables. For available data, what was the total precipitation in 2018? What was the median number of sports balls in a dumpster in 2017?

The `trash_wheels_data` includes information regarding the trash collected by Mr. Trash Wheel (water wheel vessel), which is immediately dumped into a dumpster sitting on a separate barge at the back of the device. There are `r nrow(trash_wheels_data)` observations and `r ncol(trash_wheels_data)` variables in `trash_wheel_data`. Key variables include weight_tons (weight in tons) and volume_cubic_yards (volume in cubic yards of trash in each dumpster).

The `precipitation_join` dataset includes information on rainfall in the months of 2017 and 2018. The amount of trash the device receives is highly dependent on rainfall.There are `r nrow(precipitation_join)` observations and `r ncol(precipitation_join)` in `precipitation_join`. A key variable is "total", which measures the rainfall in inches per month in a year. 

For available data, the total precipiation in 2018 was `r sum(pull(filter(precipitation_join, year == 2018), total))`.In 2017, the median number of sports balls found in a dumpster was `r median(pull(filter(trash_wheels_data, year == 2017), sports_balls))`.

## Problem 2 
## Import pols_month data 
```{r}
pols_month_data = read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(month = as.numeric(month), month = month.abb[month], month=str_to_lower(month)) %>%
  mutate(year = as.numeric(year))%>%
  mutate(president = ifelse(prez_dem, "dem", "gop")) %>%
  select(-prez_dem, -prez_gop, -day)
pols_month_data

```
## Import standard_poor_stock data 
```{r}
standard_poor_stock_data = read_csv("./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, c("month", "day", "year")) %>%
  select(year, month, -day, close) %>%
  mutate(year = as.numeric(year)) %>%
  mutate(month = as.numeric(month), month = month.abb[month], month=str_to_lower(month))
standard_poor_stock_data
```

## Import unemployment data 
```{r}
unemployment_data = read_csv("./data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer((jan:dec), names_to = "month", values_to ="percent_unemployed")
unemployment_data
```

## Merge standard_poor_stock_data and pols_month_data to form snp_pols_data 
```{r}
snp_pols_data = 
  left_join(pols_month_data, standard_poor_stock_data, by = c("year", "month"))
snp_pols_data
```
## Merge unemployment_data and snp_pols_data into snp_pols_unemp_data 
```{r}
snp_pols_unemp_data =
    left_join(snp_pols_data, unemployment_data, by = c("year", "month"))
snp_pols_unemp_data
```
## Dataset summaries 

`pols_month_data` has `r nrow(pols_month_data)` observations with `r ncol(pols_month_data)` variables. The dataset provides us with information regarding the number of national politicians who are democratic or republican at any given month ranging from `r min(pull(pols_month_data, year))` to `r max(pull(pols_month_data, year))`. Key variables include president, which tells us whether the president serving was democratic or republican in a given year and month. Other key variables include gov_gop, sen_gop, rep_gop, gov_dem, sen_dem, and rep_dem. 

`standard_poor_stock_data` has `r nrow(standard_poor_stock_data)` observations with `r ncol(standard_poor_stock_data)` variables. The dataset provides us with information related to Standard & Poor’s stock market index (S&P), ranging from `r min(pull(standard_poor_stock_data, year))` to `r max(pull(standard_poor_stock_data, year))`. A key variable is "close," which is the closing values of the S&P stock index during the associated month and year.

`unemployment_data` has `r nrow(unemployment_data)` observations with `r ncol(unemployment_data)` variables. The dataset provides us with the percentage of unemployment in the United States at a specific month and year, ranging from `r min(pull(unemployment_data, year))` to `r max(pull(unemployment_data, year))`. A key variable is percent_unemployed. 

`snp_pols_data` has `r nrow(snp_pols_data)` observations with `r ncol(snp_pols_data)` variables. The dataset joined `pols_month_data` and `standard_poor_stock_data`, providing us with information regarding the number of national politicians who are democratic or republican in a specific month and year as well as the closing values of the S&P stock index on the associated date. This dataset ranged in year from `r min(pull(snp_pols_data, year))` to `r max(pull(snp_pols_data, year))`. Key variables are: sen_gop, rep_gop, sen_dem, rep_dem, president, and close. 

`snp_pols_unemp_data` has `r nrow(snp_pols_unemp_data)` observations with `r ncol(snp_pols_unemp_data)` variables. The dataset joined `snp_pols_data` and `unemployment_data`, providing us with information regarding the number of national politicians who are democratic or republican in a specific month and year as well as the closing values of the S&P stock index and percentage of unemployment in the US on the associated date. This dataset ranged in year from `r min(pull(snp_pols_unemp_data, year))` to `r max(pull(snp_pols_unemp_data, year))`. Key variables are: sen_gop, rep_gop, sen_dem, rep_dem, president, close, and percent_unemployed.

## Problem 3 
## Import baby names data 
```{r}
baby_names_data = read_csv("./data/Popular_Baby_Names.csv") %>%
  distinct() %>%
  janitor::clean_names() %>%
  mutate(ethnicity = recode(ethnicity, "ASIAN AND PACI" = "ASIAN AND PACIFIC ISLANDER", "BLACK NON HISP" = "BLACK NON HISPANIC", "WHITE NON HISP" = "WHITE NON HISPANIC")) %>%
  mutate(ethnicity = str_to_lower(ethnicity), gender = str_to_lower(gender), childs_first_name = str_to_lower(childs_first_name))
```
## Table showing the rank in popularity of the name “Olivia” as a female baby name over time.
```{r}
baby_names_data %>%
filter(childs_first_name == "olivia") %>%
  select(-count) %>%
  pivot_wider(names_from = year_of_birth,
              values_from = rank)
```
## Table showing the most popular name among male children over time.
```{r}
baby_names_data %>%
filter(gender == "male", rank == 1) %>%
  select(-count) %>%
  pivot_wider(names_from = year_of_birth,
              values_from = rank)
```
## For male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).
```{r}
baby_names_plot = filter(baby_names_data, ethnicity == "white non hispanic", year_of_birth == "2016", gender == "male")
ggplot(baby_names_plot, aes(x = rank, y = count)) + geom_point() 
``` 
