---
title: "p8105_hw2_amh2325"
author: "Amanda Howarth"
date: "10/2/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library(tidyverse)
library(readxl)
```
## Problem 1
## Import trash_wheels data sheet 1
```{r}
trash_wheels_data =
  read_excel(path = "./data/trash_wheel.xlsx", sheet = 1) %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  select(-15, -16, -17) %>%
  mutate(sports_balls = as.integer(sports_balls))
trash_wheels_data
```
## Import precipitation_2018 data
```{r}
precipitation_2018_data = 
  read_excel(path = "./data/trash_wheel.xlsx", sheet = 6, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2018)
precipitation_2018_data
```
## Import precipitation_2017 data
```{r}
precipitation_2017_data = 
  read_excel(path = "./data/trash_wheel.xlsx", sheet = 5, range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2017)
precipitation_2017_data
```
## Merging datasets to form precipiation_join data 
```{r}
precipitation_join = 
  full_join(precipitation_2018_data, precipitation_2017_data) %>%
  mutate(month = as.numeric(month), month = month.name[month])
precipitation_join
```
Problem 1: 

## Problem 2 
## Import pols_month data 
```{r}
pols_month_data = read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(month = as.numeric(month), month = month.abb[month]) %>%
  mutate(year = as.numeric(year))%>%
  mutate(president = ifelse(prez_dem, "dem", "gop")) %>%
  select(-prez_dem, -prez_gop, -day)
pols_month_data
```

## Import standard_poor_stock data 
```{r}
standard_poor_stock_data = read_csv("./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("day", "month", "year")) %>%
  select(year, month, everything(), -day) %>%
  mutate(year = as.numeric(year))%>%
  mutate(month = as.numeric(month), month = month.abb[month])
standard_poor_stock_data
```

## Import unemployment data 
```{r}
unemployment_data = read_csv("./data/unemployment.csv") %>%
  pivot_longer((Jan:Dec), names_to = "month", values_to ="percent")%>%
  janitor::clean_names()
unemployment_data
```

## Merge standard_poor_stock_data and pols_month_data to form snp_pols_data 
```{r}
snp_pols_data = 
  left_join(pols_month_data, standard_poor_stock_data, by = c("year", "month"))
snp_pols_data
```
## Merge unemployment_data and snp_pols_data into snp_pols_unemp_data 
```{r}
snp_pols_unemp_data = 
    left_join(snp_pols_data, unemployment_data, by = c("year", "month"))
snp_pols_unemp_data
```
